A Hierarchical Bayesian Framework for Inferring the Stellar Obliquity Distribution

### 2023-05-01 src/tex/ms.tex
[showyourwork] overleaf sync

Stellar obliquity $\psi$, the angle between a planet's orbital axis and its host star's spin axis, traces the formation and evolution of a planetary system. In transiting exoplanet observations, only the sky-projected stellar obliquity $\lambda$ can be measured. To determine the true stellar obliquity $\psi$, information about the stellar inclination $i_\star$ is also needed.
In this paper, we demonstrate that while a constraint on the stellar inclination is crucial for measuring the obliquity of an individual system, it is not required for robust determination of the population-level stellar obliquity distribution. 
In practice, the constraints on the stellar obliquity distribution are mainly driven by the sky-projected stellar obliquities.
We introduce a flexible, hierarchical Bayesian framework that can be used to infer the stellar obliquity distribution solely from sky-projected stellar obliquities, including stellar inclination measurements when available.
When applying the framework to all exoplanetary systems with measured sky-projected stellar obliquity, we find that the inferred population-level obliquity distribution is unimodal and peaked at zero degrees \edits{with a fraction of $72\pm9$\,\%. The $28\pm9$\,\% fraction of misaligned systems has nearly isotropic stellar obliquities with no strong clustering near 90$\degr$.
The diverse range of stellar obliquities prefers dynamic mechanisms, such as planet-planet scattering after a convergent disk migration, which could produce both prograde and retrograde orbits with no strong inclination concentrations.}

@@

It is as yet unclear if, and to what extent, all of these physical and dynamic processes apply to exoplanetary systems. These proposed mechanisms all make different predictions on stellar obliquity distributions \citep[see][and references therein]{Albrecht22, Dawson18}. For example, the secular chaos mechanism tends to produce a stellar obliquity distribution with $\psi < 90\degr$ \citep[e.g.,][]{Teyssandier19}. \edits{The stellar von Zeipel-Lidov-Kozai (ZLK) mechanism predicts a bimodal stellar obliquity distribution, concentrated at $40\degr$ and $140\degr$ \citep[e.g.,][]{Fabrycky07, Anderson16, Vick19}, assuming a zero stellar obliquity when the ZLK oscillation begins, or a broad stellar obliquity peaked near $90\degr$ if we drop the assumption \citep{Vick23}.} The multiple-planet scattering mechanism results in a majority of aligned systems, with a small fraction of systems at a diverse range of stellar obliquities \citep[e.g.,][]{Beague12}. 
With these predictions in mind, we aim to determine the dominant mechanisms responsible for shaping close-in planetary systems by characterizing the stellar obliquity distribution of exoplanetary systems through a Bayesian approach.

@@

The relationship between the distributions of stellar obliquity, sky-projected stellar obliquity, and stellar inclination is still not fully understood. In this study, we aim to gain a deeper understanding of this relationship and develop a statistical approach to infer the stellar obliquity distribution.
\edits{In Section~\ref{sec:coords}, we find the expression of sky-projected stellar olibquity $\lambda$ and stellar inclination $i_\star$ in terms of the orbital inclination $i_{\rm orb}$, stellar obliquity $\psi$, and the azimuthal angle of the stellar spin axis $\theta$.
In Section~\ref{sec:hbm}, we introduce a flexible, hierarchical Bayesian framework that allows us to infer the stellar obliquity distribution of a sample.
In Section~\ref{sec:simulation}, we examine the framework with simulated data and show that the inferred stellar obliquity distribution from sky-projected stellar obliquities is robust even if the $i_\star$ information is not provided.
Lastly, in Section~\ref{sec:observation}, we apply the framework to real observations and derive the stellar obliquity distribution for exoplanets. We discuss the implication of the stellar obliquity distribution for Hot Jupiter origins.
}

{\color{purple}

@@

\begin{figure*}[ht!]
    \script{coordinate.py}
    \gridline{
        \fig{figures/coord_psi.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(a) The $\{\psi, \theta\}$ coordinate system. \edits{The planetary orbital axis is set to be aligned with $\vb*{z}$-axis.} The grey circle corresponds to a constant $\psi$ value and its circumference is proportional to $\sin{\psi}$.}
        \fig{figures/coord_lam.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(b) The $\{\lambda, i_\star\}$ coordinate system. \edits{The observer's light of sight is set to be aligned with $\vb*{x'}$-axis.} The grey circle corresponds to a constant $i_\star$ value and its circumference is proportional to $\sin{i_\star}$.}
    }
    \vspace*{-1cm}
    \caption{Two coordinate systems that describe the stellar spin axis $\vb{\hat{n}_\star}$ and the planet's orbital axis $\vb{\hat{n}_{\rm orb}}$. \edits{The $\{\psi, \theta\}$ coordinate system setup is motivated by the physical properties of a planetary system. The $\{\lambda, i_\star\}$ coordinate system setup is motivated by observables.} }
    \label{fig:coord}
\end{figure*}

The $\{\psi, \theta\}$ coordinate system, shown in the left panel in Figure~\ref{fig:coord}, is designed to describe the physical properties of a system. We set the planetary orbital axis $\vb{\hat{n}_{\rm orb}}$ as the $\vb{z}$-axis.
To define the stellar spin vector, we introduce the azimuthal angle of the vector around the orbital axis, $\theta$. 
The stellar spin axis $\vb{\hat{n}_\star}$ can be written in terms of $\psi$ and $\theta$ as
\begin{equation}
    \vb{\hat{n}_\star} = \sin{\psi}\cos{\theta}\vu*{x} + \sin{\psi}\sin{\theta}\vu*{y} + \cos{\psi}\vu*{z}.
\end{equation}
Particularly, we are interested in finding the distribution of the angle $\psi$.

@@

}

@@

\begin{figure}[ht!]
    \script{graph.py}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/graph.pdf}
    \caption{\edits{A probabilistic graphical model showing the conditional structure of our hierarchical Bayesian framework for inferring the stellar obliquity distribution of exoplanetary systems. The population model is parameterized by hyperparameters $\vb*{\beta}$, and it is constrained by the stellar obliquity $\psi_n$ of each $n_{\rm th}$ individual system. The stellar obliquity $\psi_n$, azimuthal angle $\theta_n$, and orbital inclination $i_{{\rm orb}, n}$ are constrained by the sky-projected stellar obliquity $\lambda_n$, and, if available, the stellar inclination $i_{\star, n}$.
    Properties of the star other than its inclination, such as its radius and rotation period, are represented by $\gamma_{\star, n}$, and these can inform the constraint on $i_{\star, n}$, if available. 
    $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity and its uncertainty, Obs$_{\star,n}$ contains all observed properties of the star with uncertainties, and $\hat{i}_{{\rm orb}, n}$ has measured orbital inclination and its uncertainty.}}
    \label{fig:graph}
\end{figure}

@@

For the stellar obliquity distribution, we model the $\cos{\psi}$ distribution instead of $\psi$ distribution to understand if the stellar obliquity is isotropically distributed. If stellar obliquity distribution is isotropic, $\cos{\psi}$ is uniformly distributed between -1 and 1.
To flexibly model the stellar obliquity distribution, \edits{we select a multi-component mixture of Beta distributions with hyperparameters $\bm{\beta} = \{\bm{w},\bm{\mu},\bm{\kappa}\}$, where each hyperparameter has a dimension of the number of components. For example, for a two-component mixture model, each hyperparameter has a dimension of 2.
We have $\bm{w} = \{w_0, w_1\}$, $\bm{\mu} = \{\mu_0, \mu_1\}$, and $\bm{\kappa} = \{\kappa_0, \kappa_1\}$.
The hyperparameter $\bm{w}$ describes the weight of each component. The $\bm{\mu}$ and 1/$\bm{\kappa}$ correspond to the mean and variance of each Beta distribution component, respectively, a reparametrization of parameters suggested by \cite{Gelman14}. The greater value of the $\kappa$, the smaller the variance (i.e., the distribution is more concentrated). The relations between the $\mu$ and $\kappa$ and the standard $\alpha$ and $\beta$ parameters in the Beta distribution are $\alpha = \mu \kappa$ and $\beta = (1-\mu) \kappa$.
This mixture distribution has the capacity to capture anything from an isotropic distribution to a strongly bimodal population.
Then, under this two-component model, the probability density function for $\cos{\psi}$ is
\begin{align}
    w_{0,1} &\sim {\rm Dirichlet}(1, 1) \nonumber\\
    u_{0,1} &\sim {\rm Beta}(\mu_{0,1}\kappa_{0,1}, (1-\mu_{0,1})\kappa_{0,1}) \nonumber\\
    \cos{\psi} &\sim 2\,(w_0 u_0 + w_1 u_1) - 1.
\end{align}
Since the Beta distribution is defined on the interval $[0, 1]$ whereas the support of $\cos{\psi}$ is from $-1$ to $1$, we extend the mixture distribution's support $[0,1]$ to $[-1,1]$ using a linear transformation (i.e., $2\,( w_0 u_0 + w_1 u_1)-1$).
For hyperparmeters $\mu$ and $\kappa$, we adopt the following priors:
\begin{align}
    \mu_{0,1} &\sim \mathcal{U}(0, 1) \nonumber\\
    \log{\kappa}_{0,1} &\sim \mathcal{N}(0, 3),
\end{align}
where $\mu_{0,1}$ is uniformly distributed between $0$ and $1$ and $\log{\kappa}_{0,1}$ is normally distributed with a mean of $0$ and standard deviation of $3$. To deal with label switching in the mixture model, we remove the symmetry by forcing the vector $\vb*{\mu}$ to be ordered. 
Notably, when applying the framework to a small sample size with $N \lesssim 50$, the choice of hyperpriors for the Beta distribution could impact the inferred distribution \citep[e.g.,][]{Nagpal22, Gelman14}. To ensure the robustness of the inferred distributions in such cases, it is crucial to test their sensitivity to different hyperpriors.
}

Next, the model parameter priors are the following:
\begin{align}
    \theta_n &\sim \mathcal{U}(0, \pi)  \nonumber\\
    \cos{i}_{{\rm orb},n} &\sim \mathcal{U}(0, 1).
\end{align}
If $\gamma_{\star, n}$ is available, we construct Normal distributions with means and standard deviations from measurements.
\edits{For the orbital inclination $i_{\rm orb}$, following the convention, we limit it to $\left[0, \pi/2\right]$. To not underestimate the $\psi$, we then set $i_\star$ to be varied from $0\degr$ to $180\degr$.
Besides, since $\lambda$ and $-\lambda$ correspond to the same $\psi$ solution, we limit $\lambda$ to $\left[0, \pi\right]$ and thus $\theta$ to $\left[0, \pi\right]$.
We find doing so greatly improves the sampling performance while not compromising the inference of the $\psi$ distribution due to the symmetry. The $\left[0, \pi\right]$ support avoids the otherwise discontinuity at $\theta = 0$ and $\theta = \pi$.}

@@

\subsection{Simulated Data}\label{subsec:sim}
To investigate the performance of our hierarchical Bayesian framework, we apply it to simulated data, generated with a known ground-truth stellar obliquity distribution.
\edits{We test the following four $\cos{\psi}$ distributions: a uniform distribution bounded between $-1$ and $1$ such that $\cos{\psi} \sim \mathcal{U}(-1,1)$, and three truncated Normal distributions bounded between $-1$ and $1$, such that $\cos{\psi} \sim \mathcal{N}(0, 0.2)$, $\cos{\psi} \sim \mathcal{N}(-0.4, 0.2)$, and $\cos{\psi} \sim \mathcal{N}(0.4, 0.2)$, where the first number is the mean of the distribution and the second number is the standard deviation.}
For each $\cos{\psi}$ distribution, we randomly generate 200 samples of sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$. We assume the stellar spin axis is uniformly distributed around the planetary orbital axis in the azimuthal direction and the orbital inclination is $90\degr$.
The sampled $i_\star$ and $\lambda$ here are \emph{true} values. 
To simulate the observation process, we add Gaussian noise to the \emph{true} $\lambda$ and $i_\star$, using uncertainties of $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$, which are typical of the literature sample \cite{Albrecht22}.
Using these simulated $\lambda$ measurements and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with both $\lambda$ and $i_\star$ likelihoods or $\lambda$ likelihood only.

In Figure~\ref{fig:simulation}, we present the results of this experiment, plotting the inferred stellar obliquity distributions, compared to the ground-truth distributions.
Since the simulated stellar obliquity distributions only have a single component, we model the data with a single Beta distribution.
Each row of Figure~\ref{fig:simulation} corresponds to a different simulation distribution. The orange curve and contours in the left column are the median, 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred $\cos{\psi}$ distribution with stellar inclination information.
The right column shows the same inferences (in blue) when constraints on stellar inclination are not included.
Surprisingly, Figure~\ref{fig:simulation} demonstrates that our inference procedure recovers the true distribution for $\cos{\psi}$ equally well, regardless of the inclusion of $i_\star$ measurements.
Despite the fact that the inferred distributions without $i_\star$ measurements have marginally wider uncertainties, as indicated by shallow color contours, the modes and widths of the inferred stellar obliquity distributions are consistent with or without $i_\star$ likelihood.
\edits{Since the injected distributions in rows 2-4 are Normal distributions, it should not be surprising that the inferred distributions, which are Beta distributions, may not exactly match the injected distributions. Given that the true distributions are not included within the support of the distributions we are using to fit, it is impressive to see how well the underlying distributions could be recovered.}

@@

Next, we aim to gain insight into the reasons behind the predominant role of the sky-projected stellar obliquity distribution and the less significant impact of the stellar inclination distribution in the inference of the stellar obliquity distribution. To simplify the problem, the orbital inclination of the transiting planet is assumed to be $90\degr$ in this illustration. As discussed earlier, the assumption will not compromise the stellar obliquity distribution inference for transiting planets.
\edits{We pair the Cartesian components in Equation~(\ref{eqn:nstar1}) and Equation~(\ref{eqn:nstar2}), assuming $i_{\rm orb} = 90\degr$, and get:
\begin{align}
    \sin{\psi}\cos{\theta} = \cos{i_\star}& \label{eq:coord1}\\
    \sin{\psi}\sin{\theta} = \sin{\lambda}\sin{i_\star}& \label{eq:coord2}\\
    \cos{\psi} = \cos{\lambda}\sin{i_\star} \label{eq:coord3}&.
\end{align}
}

First, we derive the $\lambda$ distribution for a given $\cos{\psi}$ distribution. We could find the distribution of $\cos{\lambda}$ using the Jacobian transformation from $\cos{\psi}$ and $\cos{\theta}$. Since $\psi$ and $\theta$ are assumed to be independent variables, we could marginalize over $\theta$ to find the relation between the probability density functions between $\lambda$ and $\psi$.
The Jacobian transformation follows
\begin{equation}
    p(\cos{\lambda}) = \int \abs{\pdv{\cos{\psi}}{\cos{\lambda}}} p(\cos{\psi}) p(\cos{\theta}) d\cos{\theta}.
\end{equation}
\edits{Here $p(x)$ denotes the probability density distribution of $x$.
To find $\psi$ in terms of $\lambda$ and $\theta$, we replace $\sin{i_\star}$ in Equation~(\ref{eq:coord3}) using Equation~(\ref{eq:coord1}) and find}
\begin{equation}
    \cos{\lambda} = \cos{\psi}/\sqrt{1-(1-\cos^2{\psi})\cos^2{\theta}}.
\end{equation} 
Reorganize the equation, we get 
\begin{equation}
    \cos^2{\psi} = \frac{\cos^2{\lambda}\cos^2{\theta}-\cos^2{\lambda}}{\cos^2{\lambda}\cos^2{\theta}-1}
\end{equation}
and the partial derivative 
\begin{equation}
    \abs{\pdv{\cos{\psi}}{\cos{\lambda}}} = \frac{(1-\cos^2{\theta})^{1/2}}{(1-\cos^2{\theta}\cos^2{\lambda})^{3/2}}.
\end{equation}
Since $\theta$ is uniformly distributed between $0$ and $\pi$, $p(\cos{\theta}) = p(\theta) \abs{\dv*{\theta}{\cos{\theta}}} = 1/\pi/(1-\cos^2{\theta)^{1/2}}$. Putting all the parts together, we get
\begin{equation}\label{eqn:jac_lam}
    p(\cos{\lambda}) = \frac{1}{\pi} \int_{-1}^{1} (1-\cos^2{\theta}\cos^2{\lambda})^{-3/2} p(\cos{\psi}) d\cos{\theta}.
\end{equation}
In the special case that $\cos{\psi}$ is uniformly distribution, i.e., $p(\cos{\psi}) = 1/2$, Equation~(\ref{eqn:jac_lam}) becomes $p(\cos{\lambda}) = 1/\pi/\sqrt{1-\cos^2{\lambda}}$, which is equivalent to $\lambda \sim \mathcal{U}(0, \pi)$. This suggests $\lambda$ is uniformly distributed for an isotropic $\psi$ distribution, as expected.

>>>>

Stellar obliquity, the angle between a planet's orbital axis and its host star's spin axis, traces the formation and evolution of a planetary system. In transiting exoplanet observations, only the sky-projected stellar obliquity can be measured, but this can be de-projected using an estimate of the stellar obliquity.
In this paper, we introduce a flexible, hierarchical Bayesian framework that can be used to infer the stellar obliquity distribution solely from sky-projected stellar obliquities, including stellar inclination measurements when available. 
We demonstrate that while a constraint on the stellar inclination is crucial for measuring the obliquity of an individual system, it is not required for robust determination of the population-level stellar obliquity distribution.
In practice, the constraints on the stellar obliquity distribution are mainly driven by the sky-projected stellar obliquities.

When applying the framework to all exoplanetary systems with measured sky-projected stellar obliquity, which are mostly Hot Jupiter systems, we find that the inferred population-level obliquity distribution is unimodal and peaked at zero degrees. The misaligned systems have nearly isotropic stellar obliquities with no strong clustering near 90$\degr$.
The diverse range of stellar obliquities prefers dynamic mechanisms, such as planet-planet scattering after a convergent disk migration, which could produce both prograde and retrograde orbits with no strong inclination concentrations.

@@

It is as yet unclear if, and to what extent, all of these physical and dynamic processes apply to exoplanetary systems. These proposed mechanisms all make different predictions on stellar obliquity distributions with a focus on Hot Jupiter systems \citep[see][and references therein]{Albrecht22, Dawson18}. For example, the secular chaos mechanism tends to produce a stellar obliquity distribution with $\psi < 90\degr$ \citep[e.g.,][]{Teyssandier19}. The stellar von Zeipel-Lidov-Kozai (ZLK) mechanism predicts a bimodal stellar obliquity distribution, concentrated at $40\degr$ and $140\degr$ \citep[e.g.,][]{Fabrycky07, Anderson16, Vick19}, assuming a zero stellar obliquity when the ZLK oscillation begins, or a broad stellar obliquity peaked near $90\degr$ if we drop the assumption \citep{Vick23}. The multiple-planet scattering mechanism results in a majority of aligned systems, with a small fraction of systems at a diverse range of stellar obliquities \citep[e.g.,][]{Beague12}. 
With these predictions in mind, we aim to determine the dominant mechanisms responsible for shaping close-in planetary systems by characterizing the stellar obliquity distribution of exoplanetary systems through a Bayesian approach.

@@

The relationship between the distributions of stellar obliquity, sky-projected stellar obliquity, and stellar inclination is still not fully understood. In this study, we aim to gain a deeper understanding of this relationship and develop a statistical approach to infer the stellar obliquity distribution.
In Section~\ref{sec:coords}, we find the expression of sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ in terms of the orbital inclination $i_{\rm orb}$, stellar obliquity $\psi$, and the azimuthal angle of the stellar spin axis $\theta$.
In Section~\ref{sec:hbm}, we introduce a flexible, hierarchical Bayesian framework that allows us to infer the stellar obliquity distribution of a sample.
In Section~\ref{sec:simulation}, we examine the framework with simulated data and show that the inferred stellar obliquity distribution from sky-projected stellar obliquities is robust even if the $i_\star$ information is not provided.
Lastly, in Section~\ref{sec:observation}, we apply the framework to real observations and derive the stellar obliquity distribution for exoplanets. We discuss the implication of the stellar obliquity distribution for Hot Jupiter origins.

@@

\begin{figure*}[ht!]
    \script{coordinate.py}
    \gridline{
        \fig{figures/coord_psi.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(a) The $\{\psi, \theta\}$ coordinate system. The planetary orbital axis is set to be aligned with $\vb*{z}$-axis. The grey circle corresponds to a constant $\psi$ value and its circumference is proportional to $\sin{\psi}$.}
        \fig{figures/coord_lam.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(b) The $\{\lambda, i_\star\}$ coordinate system. The observer's light of sight is set to be aligned with $\vb*{x'}$-axis. The grey circle corresponds to a constant $i_\star$ value and its circumference is proportional to $\sin{i_\star}$.}
    }
    \vspace*{-1cm}
    \caption{Two coordinate systems that describe the stellar spin axis $\vb{\hat{n}_\star}$ and the planet's orbital axis $\vb{\hat{n}_{\rm orb}}$. The $\{\psi, \theta\}$ coordinate system setup is motivated by the physical properties of a planetary system. The $\{\lambda, i_\star\}$ coordinate system setup is motivated by observables.}
    \label{fig:coord}
\end{figure*}

The $\{\psi, \theta\}$ coordinate system, shown in the left panel in Figure~\ref{fig:coord}, is designed to describe the physical properties of a system. We set the planetary orbital axis $\vb{\hat{n}_{\rm orb}}$ as the $\vb{z}$-axis.
To define the stellar spin vector, we introduce the azimuthal angle of the vector around the orbital axis, $\theta$. 
The stellar spin axis $\vb{\hat{n}_\star}$ can be written in terms of $\psi$ and $\theta$ as
\begin{equation}
    \vb{\hat{n}_\star} = \sin{\psi}\cos{\theta}\vu*{x} + \sin{\psi}\sin{\theta}\vu*{y} + \cos{\psi}\vu*{z}.
\end{equation}

@@

@@

\begin{figure}[ht!]
    \script{graph.py}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/graph.pdf}
    \caption{A probabilistic graphical model showing the conditional structure of our hierarchical Bayesian framework for inferring the stellar obliquity distribution of exoplanetary systems. The population model is parameterized by hyperparameters $\vb*{\beta}$, and it is constrained by the stellar obliquity $\psi_n$ of each $n_{\rm th}$ individual system. The stellar obliquity $\psi_n$, azimuthal angle $\theta_n$, and orbital inclination $i_{{\rm orb}, n}$ are constrained by the sky-projected stellar obliquity $\lambda_n$, and, if available, the stellar inclination $i_{\star, n}$.
    Properties of the star other than its inclination, such as its radius and rotation period, are represented by $\gamma_{\star, n}$, and these can inform the constraint on $i_{\star, n}$, if available. 
    $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity and its uncertainty, Obs$_{\star,n}$ contains all observed properties of the star with uncertainties, and $\hat{i}_{{\rm orb}, n}$ has measured orbital inclination and its uncertainty.}
    \label{fig:graph}
\end{figure}

@@

For the stellar obliquity distribution, we model the $\cos{\psi}$ distribution instead of $\psi$ distribution to understand if the stellar obliquity is isotropically distributed. If stellar obliquity distribution is isotropic, $\cos{\psi}$ is uniformly distributed between -1 and 1.
To flexibly model the stellar obliquity distribution, we select a multi-component mixture of Beta distributions with hyperparameters $\bm{\beta} = \{\bm{w},\bm{\mu},\bm{\kappa}\}$, where each hyperparameter has a dimension of the number of components. For example, for a two-component mixture model, each hyperparameter has a dimension of 2.
We have $\bm{w} = \{w_0, w_1\}$, $\bm{\mu} = \{\mu_0, \mu_1\}$, and $\bm{\kappa} = \{\kappa_0, \kappa_1\}$.
The hyperparameter $\bm{w}$ describes the weight of each component. The $\bm{\mu}$ and 1/$\bm{\kappa}$ correspond to the mean and variance of each Beta distribution component, respectively, a reparametrization of parameters suggested by \cite{Gelman14}. The greater value of the $\kappa$, the smaller the variance (i.e., the distribution is more concentrated). The relations between the $\mu$ and $\kappa$ and the standard $\alpha$ and $\beta$ parameters in the Beta distribution are $\alpha = \mu \kappa$ and $\beta = (1-\mu) \kappa$.
This mixture distribution has the capacity to capture anything from an isotropic distribution to a strongly bimodal population.
Then, under this two-component model, the probability density function for $\cos{\psi}$ is
\begin{align}
    w_{0,1} &\sim {\rm Dirichlet}(1, 1) \nonumber\\
    u_{0,1} &\sim {\rm Beta}(\mu_{0,1}\kappa_{0,1}, (1-\mu_{0,1})\kappa_{0,1}) \nonumber\\
    \cos{\psi} &\sim 2\,(w_0 u_0 + w_1 u_1) - 1.
\end{align}
Since the Beta distribution is defined on the interval $[0, 1]$ whereas the support of $\cos{\psi}$ is from $-1$ to $1$, we extend the mixture distribution's support $[0,1]$ to $[-1,1]$ using a linear transformation (i.e., $2\,( w_0 u_0 + w_1 u_1)-1$).
For hyperparmeters $\mu$ and $\kappa$, we adopt the following priors:
\begin{align}
    \mu_{0,1} &\sim \mathcal{U}(0, 1) \nonumber\\
    \log{\kappa}_{0,1} &\sim \mathcal{N}(0, 3),
\end{align}
where $\mu_{0,1}$ is uniformly distributed between $0$ and $1$ and $\log{\kappa}_{0,1}$ is normally distributed with a mean of $0$ and standard deviation of $3$. To deal with label switching in the mixture model, we remove the symmetry by forcing the vector $\vb*{\mu}$ to be ordered. 
Notably, when applying the framework to a small sample size with $N \lesssim 50$, the choice of hyperpriors for the Beta distribution could impact the inferred distribution \citep[e.g.,][]{Nagpal22, Gelman14}. To ensure the robustness of the inferred distributions in such cases, it is crucial to test their sensitivity to different hyperpriors.

Next, the model parameter priors are the following:
\begin{align}
    \theta_n &\sim \mathcal{U}(0, \pi)  \nonumber\\
    \cos{i}_{{\rm orb},n} &\sim \mathcal{U}(0, 1).
\end{align}
If $\gamma_{\star, n}$ is available, we construct Normal distributions with means and standard deviations from measurements.
For the orbital inclination $i_{\rm orb}$, following the convention, we limit it to $\left[0, \pi/2\right]$. To not underestimate the $\psi$, we then set $i_\star$ to be varied from $0\degr$ to $180\degr$.
Besides, since $\lambda$ and $-\lambda$ correspond to the same $\psi$ solution, we limit $\lambda$ to $\left[0, \pi\right]$ and thus $\theta$ to $\left[0, \pi\right]$.
We find doing so greatly improves the sampling performance while not compromising the inference of the $\psi$ distribution due to the symmetry. The $\left[0, \pi\right]$ support avoids the otherwise discontinuity at $\theta = 0$ and $\theta = \pi$.

@@

\subsection{Simulated Data}\label{subsec:sim}
To investigate the performance of our hierarchical Bayesian framework, we apply it to simulated data, generated with a known ground-truth stellar obliquity distribution.
We test the following four $\cos{\psi}$ distributions: a uniform distribution bounded between $-1$ and $1$ such that $\cos{\psi} \sim \mathcal{U}(-1,1)$, and three truncated Normal distributions bounded between $-1$ and $1$, such that $\cos{\psi} \sim \mathcal{N}(0, 0.2)$, $\cos{\psi} \sim \mathcal{N}(-0.4, 0.2)$, and $\cos{\psi} \sim \mathcal{N}(0.4, 0.2)$, where the first number is the mean of the distribution and the second number is the standard deviation.
For each $\cos{\psi}$ distribution, we randomly generate 200 samples of sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$. We assume the stellar spin axis is uniformly distributed around the planetary orbital axis in the azimuthal direction and the orbital inclination is $90\degr$.
The sampled $i_\star$ and $\lambda$ here are \emph{true} values. 
To simulate the observation process, we add Gaussian noise to the \emph{true} $\lambda$ and $i_\star$, using uncertainties of $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$, which are typical of the literature sample \cite{Albrecht22}.
Using these simulated $\lambda$ measurements and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with both $\lambda$ and $i_\star$ likelihoods or $\lambda$ likelihood only.

In Figure~\ref{fig:simulation}, we present the results of this experiment, plotting the inferred stellar obliquity distributions, compared to the ground-truth distributions.
Since the simulated stellar obliquity distributions only have a single component, we model the data with a single Beta distribution.
Each row of Figure~\ref{fig:simulation} corresponds to a different simulation distribution. The orange curve and contours in the left column are the median, 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred $\cos{\psi}$ distribution with stellar inclination information.
The right column shows the same inferences (in blue) when constraints on stellar inclination are not included.
Surprisingly, Figure~\ref{fig:simulation} demonstrates that our inference procedure recovers the true distribution for $\cos{\psi}$ equally well, regardless of the inclusion of $i_\star$ measurements.
Despite the fact that the inferred distributions without $i_\star$ measurements have marginally wider uncertainties, as indicated by shallow color contours, the modes and widths of the inferred stellar obliquity distributions are consistent with or without $i_\star$ likelihood.
Since the injected distributions in rows 2-4 are Normal distributions, it should not be surprising that the inferred distributions, which are Beta distributions, may not exactly match the injected distributions. Given that the true distributions are not included within the support of the distributions we are using to fit, it is impressive to see how well the underlying distributions could be recovered.

@@

Next, we aim to gain insight into the reasons behind the predominant role of the sky-projected stellar obliquity distribution and the less significant impact of the stellar inclination distribution in the inference of the stellar obliquity distribution. To simplify the problem, the orbital inclination of the transiting planet is assumed to be $90\degr$ in this illustration. As discussed earlier, the assumption will not compromise the stellar obliquity distribution inference for transiting planets.
We pair the Cartesian components in Equation~(\ref{eqn:nstar1}) and Equation~(\ref{eqn:nstar2}), assuming $i_{\rm orb} = 90\degr$, and get:
\begin{align}
    \sin{\psi}\cos{\theta} = \cos{i_\star}& \label{eq:coord1}\\
    \sin{\psi}\sin{\theta} = \sin{\lambda}\sin{i_\star}& \label{eq:coord2}\\
    \cos{\psi} = \cos{\lambda}\sin{i_\star} \label{eq:coord3}&.
\end{align}

First, we derive the $\lambda$ distribution for a given $\cos{\psi}$ distribution. We could find the distribution of $\cos{\lambda}$ using the Jacobian transformation from $\cos{\psi}$ and $\cos{\theta}$. Since $\psi$ and $\theta$ are assumed to be independent variables, we could marginalize over $\theta$ to find the relation between the probability density functions between $\lambda$ and $\psi$.
The Jacobian transformation follows
\begin{equation}
    p(\cos{\lambda}) = \int \abs{\pdv{\cos{\psi}}{\cos{\lambda}}} p(\cos{\psi}) p(\cos{\theta}) d\cos{\theta}.
\end{equation}
Here $p(x)$ denotes the probability density distribution of $x$.
To find $\psi$ in terms of $\lambda$ and $\theta$, we replace $\sin{i_\star}$ in Equation~(\ref{eq:coord3}) using Equation~(\ref{eq:coord1}) and find
\begin{equation}
    \cos{\lambda} = \frac{\cos{\psi}}{\sqrt{1-(1-\cos^2{\psi})\cos^2{\theta}}}.
\end{equation} 
Reorganize the equation, we get 
\begin{equation}
    \cos^2{\psi} = \frac{\cos^2{\lambda}\cos^2{\theta}-\cos^2{\lambda}}{\cos^2{\lambda}\cos^2{\theta}-1}
\end{equation}
and the partial derivative 
\begin{equation}
    \abs{\pdv{\cos{\psi}}{\cos{\lambda}}} = \frac{(1-\cos^2{\theta})^{1/2}}{(1-\cos^2{\theta}\cos^2{\lambda})^{3/2}}.
\end{equation}
Since $\theta$ is uniformly distributed between $0$ and $\pi$, $p(\cos{\theta}) = p(\theta) \abs{\dv*{\theta}{\cos{\theta}}} = 1/\pi/(1-\cos^2{\theta)^{1/2}}$. Putting all the parts together, we get
\begin{equation}\label{eqn:jac_lam}
    p(\cos{\lambda}) = \frac{1}{\pi} \int_{-1}^{1} (1-\cos^2{\theta}\cos^2{\lambda})^{-3/2} p(\cos{\psi}) d\cos{\theta}.
\end{equation}
In the special case that $\cos{\psi}$ is uniformly distribution, i.e., $p(\cos{\psi}) = 1/2$, Equation~(\ref{eqn:jac_lam}) becomes $p(\cos{\lambda}) = 1/\pi/\sqrt{1-\cos^2{\lambda}}$, which is equivalent to $\lambda \sim \mathcal{U}(0, \pi)$. This suggests $\lambda$ is uniformly distributed for an isotropic $\psi$ distribution, as expected.
